\vspace{-\baselineskip} Fast and scalable Gaussian process modeling with applications to astronomical time series \vspace{-2\baselineskip}

### 2017-06-02 paper/ms.tex
words about rotation and turning off omp for mkl

% Matrix fix:
% http://tex.stackexchange.com/questions/317824/letter-c-appearing-inside-pmatrix-environment-with-aastex
\makeatletter
\def\env@matrix{\hskip -\arraycolsep % taken from amsmath.sty lines 895ff
  \let\@ifnextchar\new@ifnextchar
  \array{*{\c@MaxMatrixCols}c}}
\makeatother

@@

% Projects:
\newcommand{\project}[1]{\textsf{#1}}
\newcommand{\kepler}{\project{Kepler}}
\newcommand{\tess}{\project{TESS}}
\newcommand{\celerite}{\project{celerite}}
\newcommand{\celeriteterm}{\emph{celerite}}
\newcommand{\emcee}{\project{emcee}}

@@

@@

In this paper, we present a method for computing a class of GP models that
scales linearly with the number of data points $\mathcal{O}(N)$ for one
dimensional data sets.
\dfmtodo{This method is a generalization of a method developed by
\citet{Ambikasaran:2015}
 ({\bf This is no longer correct -different Ambikasaran result!})}
that was, in turn, built on intuition from a twenty
year old paper \citep{Rybicki:1995}.
This method can only be used with one-dimensional datasets and the covariance
function must be represented by a mixture of exponentials; we will return to a
discussion of what this means in detail in the following sections.
However, there is no further constraint on the data or the model.
In particular, the measurements don't need to be evenly spaced and the
uncertainties can be heteroscedastic (non-uniform).

@@

To demonstrate the application of \celerite, we use it to perform posterior
inference in two examples with realistic simulated data and three examples of
real-world research problems.

start by inferring
posterior constraints on the parameters of a GP model applied to several
simulated datasets with known properties.

In the following section, we expand on these examples by applying
\celerite\ to real datasets.
In the first example, we demonstrate that \celerite\ can be used to infer
the power spectrum of a process when the data are generated from a
\celeriteterm\ model.
In the second example, we demonstrate that \celerite\ can be used as an
effective model even if the true process cannot be represented in the space of
allowed models.
This is an interesting example because, when analyzing real data, we rarely
have any fundamental reason to believe that the data were generated by a GP
model with a specific kernel.
Even in these cases, GPs can be useful effective models and \celerite\
provides computational advantages over other GP methods.

In this section, we demonstrate several use cases of \celerite\ when
applied to real datasets.
Each of these examples touches on an active area of research so we limit our
examples to be qualitative in nature and do not claim that \celerite\ is the
optimum method, but we hope these examples encourage interested readers to
investigate the applicability of \celerite\ to their research.

All of the following examples show time domain datasets with a clear bias in
favor of large homogeneous photometric surveys, but these methods can
similarly be applied to spectroscopy, where wavelength~--~instead of
time~--~is the independent coordinate and other one-dimensional domains
\citep[see][for a potential application]{Czekala:2017}.

\response{In each example, the joint posterior probability density is given by
\begin{eqnarray}\eqlabel{joint-post}
p(\bvec{\theta},\,\bvec{\alpha}\,|\,\bvec{y},\,X) \propto
p(\bvec{y}\,|\,{X,\,\bvec{\theta}},\,\bvec{\alpha})\,
p(\bvec{\theta},\,\bvec{\alpha})
\end{eqnarray}
where $p(\bvec{y}\,|\,{X,\,\bvec{\theta}},\,\bvec{\alpha})$ is the GP
likelihood defined in \eq{gp-likelihood} and
$p(\bvec{\theta},\,\bvec{\alpha})$ is the joint prior probability density for
the parameters.
The specific parameters and priors are discussed in detail in each section
but, in most cases, we assume separable uniform priors with plausibly broad
support.
}

\subsection{Recovery of a celerite process}

In this first example, we simulate a dataset using a known \celeriteterm\
process and fit it with \celerite\ to demonstrate that valid inferences can be
made in this idealized case.
\response{We simulated a small ($N = \exampleindata$) dataset using a GP model
with a SHO kernel (\eqalt{sho-kernel}) with parameters $S_0 =
1\,\mathrm{ppm}^2$, $\omega_0 = e^2\,\mathrm{day}^{-1}$, and $Q = e^2$, where
the units are arbitrarily chosen to simplify the discussion.
We add further white noise with the amplitude $\sigma_n = 2.5\, \mathrm{ppm}$
to each data point.
The simulated data are plotted in the top left panel of
\Figure{simulated-correct}.

@@

We then sample the joint posterior probability (\eqalt{joint-post}) using
\emcee\ \citep{Goodman:2010, Foreman-Mackey:2013}.
We initialize \exampleinwalkers~walkers in a Gaussian ball with width
$10^{-4}$ in each parameter around the MAP parameter vector, run
\exampleinburn~steps of burn-in, and \exampleinsteps~steps of MCMC.
To asses convergence, we estimate the mean integrated autocorrelation time of
the chain across parameters \citep{Sokal:1989, Goodman:2010} and find that the
chain results in \exampleineff~effective samples.

@@

\newpage
\subsection{Inferences with the ``wrong'' model}

@@

The data are simulated from a quasiperiodic GP with the kernel
\begin{eqnarray}\eqlabel{sim-wrong-true}
k_\mathrm{true} (\tau) = \alpha\,
    \exp\left(-\frac{\tau^2}{2\,\lambda^2}\right)\,
    \cos\left(\frac{2\,\pi\,\tau}{P_\mathrm{true}}\right)
\end{eqnarray}
where $P_\mathrm{true}$ is the fundamental period of the process.
This autocorrelation structure corresponds to the power spectrum
\citep{Wilson:2013}
\begin{eqnarray}
S_\mathrm{true} (\omega) = \frac{\lambda\,\alpha}{2}\,\left[
    \exp\left(-\frac{\lambda^2}{2}\,\left(\omega-
        \frac{2\,\pi}{P_\mathrm{true}}\right)^2\right) +
    \exp\left(-\frac{\lambda^2}{2}\,\left(\omega+
        \frac{2\,\pi}{P_\mathrm{true}}\right)^2\right)
\right]
\end{eqnarray}
which, for large values of $\omega$, falls off exponentially. \dfmtodo{$\alpha$
means the parameter vector in definition of likelihood.}
When compared to \eq{celerite-psd}~--~which, for large $\omega$, goes as
$\omega^{-4}$ at most~--~it is clear that a \celeriteterm\ model can never
exactly reproduce the structure of this process.
That being said, we demonstrate that robust inferences can be made about
$P_\mathrm{true}$ even with this effective model.

\response{We generate a realization of a GP model with the kernel given in
\eq{sim-wrong-true} with $N = \exampleiindata$, $\alpha = 1\,\mathrm{ppm}^2$,
$\lambda = 5\,\mathrm{day}$, and $P_\mathrm{true} = 1\,\mathrm{day}$.
We also add white noise with amplitude $\sigma = 0.5\,\mathrm{ppm}$ to each
data point.
The top left panel of \Figure{simulated-wrong} shows this simulated dataset.

We then fit this simulated data using the product of two SHO terms
(\eqalt{sho-kernel}) where one of the terms has $S_0 = 1$ and $Q =
1/\sqrt{2}$ fixed.
The kernel for this model is
\begin{eqnarray}
k(\tau) = k_\mathrm{SHO}(\tau;\,S_0,\,Q,\,\omega_1) \,
    k_\mathrm{SHO}(\tau;\,S_0 = 1,\,Q = 1/\sqrt{2},\,\omega_2)
\end{eqnarray}
where $k_\mathrm{SHO}$ is defined in \eq{sho-kernel} and the inferred period
is related to the parameters by $P=2\,\pi/\omega_1$.
We note that using \eq{product-rule}, the product of two \celeriteterm\ terms
can also be expressed using \celerite.

@@

For comparison, we run the same number of steps of MCMC to sample the
``correct'' joint posterior density.
For this reference inference, we use a GP likelihood with the kernel given by
\eq{sim-wrong-true} and choose log-uniform priors on each of the
three parameters $\ln \alpha/\mathrm{ppm}^2 \sim \mathcal{U}(-10,\,10)$,
$\ln \lambda/\mathrm{day} \sim \mathcal{U}(-10,\,10)$, and
$\ln P_\mathrm{true}/\mathrm{day} \sim \mathcal{U}(-10,\,10)$.

@@

\subsection{Stellar rotation} \sectlabel{rotation}

A source of variability that can be measured from time series measurements of
stars is rotation.
The inhomogeneous surface of the star (spots, plage, \etc) imprints itself as
quasiperiodic variations in photometric or spectroscopic observations
\citep{Dumusque:2014}.
It has been demonstrated that for light curves with nearly uniform sampling,
the empirical autocorrelation function provides a reliable estimate of the
rotation period of a star \citep{Mcquillan:2013, Mcquillan:2014, Aigrain:2015}
and that a GP model with a
quasiperiodic covariance function can be used to make probabilistic
measurements even with sparsely sampled data (R.~Angus, \etal\ submitted).
The covariance function used for this type of analysis has the form
\begin{eqnarray}\eqlabel{sine2}
k(\tau) = A\,\exp\left(-\frac{\tau^2}{2\,\ell^2} -
    \Gamma\,\sin^2\left(\frac{\pi\,\tau}{P} \right) \right)
\end{eqnarray}
where $P$ is the rotation period of the star.
GP modeling with the same kernel function has been proposed as a method of
measuring the mean periodicity in quasiperiodic photometric time series in
general \citep{Wang:2012}.
The key difference between \eq{sine2} and other quasiperiodic kernels is that
it is positive for all values of $\tau$.
We can construct a simple \celeriteterm\ covariance function with similar properties
as follows
\begin{eqnarray}\eqlabel{rot-kernel}
k(\tau) = \frac{a}{2+b}\,e^{-\tau/c}\,\left[
    \cos\left(\frac{2\,\pi\,\tau}{P}\right) + (1 + b)
\right]
\end{eqnarray}
for $a>0$, $b>0$, and $c>0$. \dfmtodo{a,b,c already have different meaning
in celerite kernel.}
The covariance function in \eq{rot-kernel} cannot exactly reproduce \eq{sine2}
but, since \eq{sine2} is only an effective model, \eq{rot-kernel} can be used
as a drop-in replacement for a substantial gain in computational efficiency.

To demonstrate this method, we fit a \celeriteterm\ model with a kernel given
by \eq{rot-kernel} to a \kepler\ light curve for the star KIC~1430163.
This star has a published rotation period of $3.88 \pm 0.58\,\mathrm{days}$,
measured using traditional periodogram and autocorrelation function approaches
applied to \kepler\ data from Quarters 0--16 \citep{Mathur:2014}.
We used \emcee\ to sample the joint posterior probability density for the four
parameters in \eq{rot-kernel}, conditioned on two quarters of \kepler\ data,
and find a constraint on the rotation period of $P = \rotationperiod$.
This is in good agreement with the literature value with somewhat smaller
uncertainties.
A detailed comparison of GP rotation period measurements and the traditional
methods is beyond the scope of this paper, but forthcoming work (R.~Angus,
\etal, submitted) demonstrates that GP inferences are, at a population level,
more reliable than other methods.
The left panels of \Figure{rotation} show a subset of the data used in this
example and the residuals away from the maximum likelihood predictive mean.
The right panel of \Figure{rotation} shows the marginalized posterior
distribution for the rotation period.

@@

\begin{figure}[p]
\begin{center}
\includegraphics[width=0.8\textwidth]{figures/rotation/rotation.pdf}
\caption{Inferred constraints on a quasiperiodic GP model using the covariance
    function in \eq{rot-kernel} and two quarters of \kepler\ data.
(top left) The \kepler\ data (black points) and the maximum likelihood model
    prediction (blue curve) for a 60~day subset of the data used.
    The solid blue line shows the predictive mean and the blue contours show
    the predictive standard deviation.
(bottom left) The residuals between the mean predictive model and the data
    shown in the top left figure.
(right) The posterior constraint on the rotation period of KIC~1430163 using
    the dataset and model from \Figure{rotation}.
    The period is the parameter $P$ in \eq{rot-kernel} and this figure shows
    the posterior distribution marginalized over all other nuisance parameters
    in \eq{rot-kernel}.
    This is consistent with the published rotation period made using the full
    \kepler\ baseline \citep{Mathur:2014}.
    The error bars from the literature are indicated by the two vertical
    dashed lines.
    \figurelabel{rotation}}
\end{center}
\end{figure}

>>>>

% Matrix fix:
% http://tex.stackexchange.com/questions/317824/letter-c-appearing-inside-pmatrix-environment-with-aastex
\makeatletter
\def\env@matrix{\hskip -\arraycolsep\let\@ifnextchar\new@ifnextchar\array{*{\c@MaxMatrixCols}c}}
\makeatother

@@

% Projects:
\newcommand{\project}[1]{\textsf{#1}}
\newcommand{\kepler}{\project{Kepler}}
\newcommand{\lsst}{\project{LSST}}
\newcommand{\tess}{\project{TESS}}
\newcommand{\celerite}{\project{celerite}}
\newcommand{\celeriteterm}{\emph{celerite}}
\newcommand{\emcee}{\project{emcee}}

@@

\dfmtodo{Talk about LSST here too!}

@@

In this paper, we present a method for computing a class of GP models that
scales linearly with the number of data points $\mathcal{O}(N)$ for one
dimensional data sets.
\dfmtodo{This method is a generalization of a method developed by
\citet{Ambikasaran:2015}
 ({\bf This is no longer correct -different Ambikasaran result!})}
that was, in turn, built on intuition from a twenty
year old paper \citep{Rybicki:1995}.
This method can only be used with one-dimensional datasets and the covariance
function must be represented by a mixture of exponentials; we will return to a
discussion of what this means in detail in the following sections.
However, there is no further constraint on the data or the model.
In particular, the measurements don't need to be evenly spaced and the
uncertainties can be heteroscedastic.

@@

\response{
To show the application of \celerite, we use it to perform posterior
inference in two examples with realistic simulated data and three examples of
real-world research problems.
These examples have been chosen to justify the application of \celerite\ for a
range of use cases, but they are by no means exhaustive.
These examples are framed in the time domain with a clear bias in favor of
large homogeneous photometric surveys, but \celerite\ can also be used for
other one-dimensional problems like, for example, spectroscopy, where
wavelength~--~instead of time~--~is the independent coordinate \citep[see][for
a potential application]{Czekala:2017}.

In the first example, we demonstrate that \celerite\ can be used to infer
the power spectrum of a process when the data are generated from a
\celeriteterm\ model.
In the second example, we demonstrate that \celerite\ can be used as an
effective model even if the true process cannot be represented in the space of
allowed models.
This is an interesting example because, when analyzing real data, we rarely
have any fundamental reason to believe that the data were generated by a GP
model with a specific kernel.
Even in these cases, GPs can be useful effective models and \celerite\
provides computational advantages over other GP methods.

The next three examples show \celerite\ used to infer the properties of stars
and exoplanets observed by NASA's \kepler\ Mission.
Each of these examples touches on an active area of research so we limit our
examples to be qualitative in nature and do not claim that \celerite\ is the
optimal method, but we hope these examples encourage interested readers to
investigate the applicability of \celerite\ to their research.

In each example, the joint posterior probability density is given by
\begin{eqnarray}\eqlabel{joint-post}
p(\bvec{\theta},\,\bvec{\alpha}\,|\,\bvec{y},\,X) \propto
p(\bvec{y}\,|\,{X,\,\bvec{\theta}},\,\bvec{\alpha})\,
p(\bvec{\theta},\,\bvec{\alpha})
\end{eqnarray}
where $p(\bvec{y}\,|\,{X,\,\bvec{\theta}},\,\bvec{\alpha})$ is the GP
likelihood defined in \eq{gp-likelihood} and $p(\bvec{\theta},\,\bvec{\alpha})$
is the joint prior probability density for the parameters.
We sample each posterior density using MCMC and investigate the performance of
the model when used to interpret the physical processes that generated the
data.
The specific parameters and priors are discussed in detail in each section,
but we generally assume separable uniform priors with plausibly broad support.
}

\subsection{Example 1: Recovery of a celerite process}

In this first example, we simulate a dataset using a known \celeriteterm\
process and fit it with \celerite\ to show that valid inferences can be made
in this idealized case.
\response{We simulated a small ($N = \exampleindata$) dataset using a GP model
with a SHO kernel (\eqalt{sho-kernel}) with parameters $S_0 =
1\,\mathrm{ppm}^2$, $\omega_0 = e^2\,\mathrm{day}^{-1}$, and $Q = e^2$, where
the units are arbitrarily chosen to simplify the discussion.
We add further white noise with the amplitude $\sigma_n = 2.5\, \mathrm{ppm}$
to each data point.
The simulated data are plotted in the top left panel of
\Figure{simulated-correct}.

@@

We then sample the joint posterior probability (\eqalt{joint-post}) using
\emcee\ \citep{Goodman:2010, Foreman-Mackey:2013}.
We initialize \exampleinwalkers~walkers in a Gaussian ball with width
$10^{-4}$ in each parameter around the MAP parameter vector, run
\exampleinburn~steps of burn-in, and \exampleinsteps~steps of MCMC.
To assess convergence, we estimate the mean integrated autocorrelation time of
the chain across parameters \citep{Sokal:1989, Goodman:2010} and find that the
chain results in \exampleineff~effective samples.

@@

\subsection{Example 2: Inferences with the ``wrong'' model}

@@

The data are simulated from a quasiperiodic GP with the kernel
\begin{eqnarray}\eqlabel{sim-wrong-true}
k_\mathrm{true} (\tau) = A\,
    \exp\left(-\frac{\tau^2}{2\,\lambda^2}\right)\,
    \cos\left(\frac{2\,\pi\,\tau}{P_\mathrm{true}}\right)
\end{eqnarray}
where $P_\mathrm{true}$ is the fundamental period of the process.
This autocorrelation structure corresponds to the power spectrum
\citep{Wilson:2013}
\begin{eqnarray}
S_\mathrm{true} (\omega) = \frac{\lambda\,A}{2}\,\left[
    \exp\left(-\frac{\lambda^2}{2}\,\left(\omega-
        \frac{2\,\pi}{P_\mathrm{true}}\right)^2\right) +
    \exp\left(-\frac{\lambda^2}{2}\,\left(\omega+
        \frac{2\,\pi}{P_\mathrm{true}}\right)^2\right)
\right]
\end{eqnarray}
which, for large values of $\omega$, falls off exponentially.
When compared to \eq{celerite-psd}~--~which, for large $\omega$, goes as
$\omega^{-4}$ at most~--~it is clear that a \celeriteterm\ model can never
exactly reproduce the structure of this process.
That being said, we demonstrate that robust inferences can be made about
$P_\mathrm{true}$ even with this effective model.

\response{We generate a realization of a GP model with the kernel given in
\eq{sim-wrong-true} with $N = \exampleiindata$, $A = 1\,\mathrm{ppm}^2$,
$\lambda = 5\,\mathrm{day}$, and $P_\mathrm{true} = 1\,\mathrm{day}$.
We also add white noise with amplitude $\sigma = 0.5\,\mathrm{ppm}$ to each
data point.
The top left panel of \Figure{simulated-wrong} shows this simulated dataset.

We then fit this simulated data using the product of two SHO terms
(\eqalt{sho-kernel}) where one of the terms has $S_0 = 1$ and $Q =
1/\sqrt{2}$ fixed.
The kernel for this model is
\begin{eqnarray}
k(\tau) = k_\mathrm{SHO}(\tau;\,S_0,\,Q,\,\omega_1) \,
    k_\mathrm{SHO}(\tau;\,S_0 = 1,\,Q = 1/\sqrt{2},\,\omega_2)
\end{eqnarray}
where $k_\mathrm{SHO}$ is defined in \eq{sho-kernel} and the period of the
process is $P=2\,\pi/\omega_1$.
We note that using \eq{product-rule}, the product of two \celeriteterm\ terms
can also be expressed using \celerite.

@@

For comparison, we run the same number of steps of MCMC to sample the
``correct'' joint posterior density.
For this reference inference, we use a GP likelihood with the kernel given by
\eq{sim-wrong-true} and choose log-uniform priors on each of the
three parameters $\ln A/\mathrm{ppm}^2 \sim \mathcal{U}(-10,\,10)$,
$\ln \lambda/\mathrm{day} \sim \mathcal{U}(-10,\,10)$, and
$\ln P_\mathrm{true}/\mathrm{day} \sim \mathcal{U}(-10,\,10)$.

@@

\subsection{Example 3: Stellar rotation} \sectlabel{rotation}

A source of variability that can be measured from time series measurements of
stars is rotation.
The inhomogeneous surface of the star (spots, plage, \etc) imprints itself as
quasiperiodic variations in photometric or spectroscopic observations
\citep{Dumusque:2014}.
It has been demonstrated that for light curves with nearly uniform sampling,
the empirical autocorrelation function provides a reliable estimate of the
rotation period of a star \citep{Mcquillan:2013, Mcquillan:2014, Aigrain:2015}
and that a GP model with a
quasiperiodic covariance function can be used to make probabilistic
measurements even with sparsely sampled data (R.~Angus, \etal\ submitted).
The covariance function used for this type of analysis has the form
\begin{eqnarray}\eqlabel{sine2}
k(\tau) = A\,\exp\left(-\frac{\tau^2}{2\,\ell^2} -
    \Gamma\,\sin^2\left(\frac{\pi\,\tau}{P} \right) \right)
\end{eqnarray}
where $P$ is the rotation period of the star.
GP modeling with the same kernel function has been proposed as a method of
measuring the mean periodicity in quasiperiodic photometric time series in
general \citep{Wang:2012}.
The key difference between \eq{sine2} and other quasiperiodic kernels is that
it is positive for all values of $\tau$.
We construct a simple \celeriteterm\ covariance function with similar
properties as follows
\begin{eqnarray}\eqlabel{rot-kernel}
k(\tau) = \frac{B}{2+C}\,e^{-\tau/L}\,\left[
    \cos\left(\frac{2\,\pi\,\tau}{P}\right) + (1 + C)
\right]
\end{eqnarray}
for $B>0$, $C>0$, and $L>0$.
The covariance function in \eq{rot-kernel} cannot exactly reproduce \eq{sine2}
but, since \eq{sine2} is only an effective model, \eq{rot-kernel} can be used
as a drop-in replacement for a significant gain in computational efficiency.

\response{GPs have been used to measure stellar rotation periods for
individual datasets \citep[for example][]{Littlefair:2017}, but the
computational cost of traditional GP methods has hindered the industrial
application to existing surveys like \kepler\ with hundreds of thousands of
targets.
The increase in computational efficiency and scalability provided by
\celerite\ opens the possibility of inferring rotation periods using GPs at
scale of existing and forthcoming surveys like \kepler, \tess, and \lsst.
}

As a demonstration, we fit a \celeriteterm\ model with a kernel given
by \eq{rot-kernel} to a \kepler\ light curve for the star KIC~1430163.
This star has a published rotation period of $3.88 \pm 0.58\,\mathrm{days}$,
measured using traditional periodogram and autocorrelation function approaches
applied to \kepler\ data from Quarters 0--16 \citep{Mathur:2014}.

\response{We select about 180~days of contiguous observations of
KIC~1430163 from \kepler.
This dataset has \exampleiiindata\ measurements and, using a tuned linear
algebra
implementation\footnote{We use the Intel Math Kernel Library
\url{https://software.intel.com/en-us/intel-mkl}}, a
single evaluation of the likelihood requires over 8~seconds on a modern Intel
CPU.
This calculation, using \celerite\ with the model in \eq{rot-kernel}, only
takes $\sim 1.5\,\mathrm{ms}$~--~a speed-up of more than three orders of
magnitude per model evaluation.

We set the mean function $\mu_\bvec{\theta}$ to zero and the remaining
parameters $\bvec{\alpha}$ and their priors are listed in
Table~\ref{tab:example-3-params}.
As with the earlier examples, we start by estimating the MAP parameters using
\project{L-BFGS-B} and initialize \exampleiiinwalkers~walkers by sampling from
an isotropic Gaussian with a standard deviation of $10^{-5}$ centered on the
MAP parameters.
}

We used \emcee\ to sample the joint posterior probability density for the four
parameters in \eq{rot-kernel}, conditioned on two quarters of \kepler\ data,
and find a constraint on the rotation period of $P = \rotationperiod$.
This is in good agreement with the literature value with smaller
uncertainties.
A detailed comparison of GP rotation period measurements and the traditional
methods is beyond the scope of this paper, but forthcoming work (R.~Angus,
\etal, submitted) demonstrates that GP inferences are, at a population level,
more reliable than other methods.
The left panels of \Figure{rotation} show a subset of the data used in this
example and the residuals away from the maximum likelihood predictive mean.
The right panel of \Figure{rotation} shows the marginalized posterior
distribution for the rotation period.

@@

\begin{figure}[p]
\begin{center}
\includegraphics[width=0.8\textwidth]{figures/rotation/rotation.pdf}
\caption{Inferred constraints on a quasiperiodic GP model using the covariance
    function in \eq{rot-kernel} and two quarters of \kepler\ data.
(top left) The \kepler\ data (black points) and the maximum likelihood model
    prediction (blue curve) for a 60~day subset of the data used.
    The solid blue line shows the predictive mean and the blue contours show
    the predictive standard deviation.
(bottom left) The residuals between the mean predictive model and the data
    shown in the top left figure.
(right) The posterior constraint on the rotation period of KIC~1430163 using
    the dataset and model from \Figure{rotation}.
    The period is the parameter $P$ in \eq{rot-kernel} and this figure shows
    the posterior distribution marginalized over all other nuisance parameters
    in \eq{rot-kernel}.
    This is consistent with the published rotation period made using the full
    \kepler\ baseline shown as a vertical gray line \citep{Mathur:2014}.
    \figurelabel{rotation}}
\end{center}
\end{figure}
